Interspecies comparisons of Mg/Ca ratios in limpet shells

### 2024-06-14 Manuscript.qmd
Added ORK-LT5

```{r Libray and datasets}
#| echo: FALSE
#| message: FALSE
#| warning: FALSE

@@

# Scripts
source("DTW_Script.R")

@@

@@

@@

sample <- Samples[6]

@@

  temp_MgCa <- MgCa %>% 
    filter(sampleID == sample) %>% 
    filter(!(sampleID=="ND-1016-4"&dist<0.3)) %>% 
    select(mg_ca) %>% 
    mutate(sample_seq = row_number())

@@

P_all <- Corr_data %>% 
  pivot_wider(values_from = "value",names_from = "proxy") %>% 
  mutate(region = substr(sampleID, 1, 1)) %>% 
  ggplot()+
  aes(d18o,mg_ca,col=sampleID)+
  geom_point(show.legend = FALSE)+
  labs(x=expression("‰ δ" ^ "18" * "O [VPDB]"),y="Mg/Ca intensity ratio")+
  stat_smooth(aes(fill = sampleID,col=sampleID), method = "lm", formula = y~x,show.legend = FALSE) +
  # stat_regline_equation(label.x = 2, #label.y = 32
  #   aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
  #   formula = y~x
  # ) +
  scale_colour_manual(values = c( "#33a02c","#2ecc71", "#ffc100", "#f39c12", "#007f7f", "#1f78b4", "#99d8c9", "#542788","black", "#d73027", "#EC7063")) +
scale_fill_manual(values = c( "#33a02c","#2ecc71", "#ffc100", "#f39c12", "#007f7f", "#1f78b4", "#99d8c9", "#542788","black", "#d73027", "#EC7063"))#+
  # facet_wrap(~region,ncol=2,scales = "free")

@@

@@

|-------------------|---------------|---------------|---------------|---------------|

@@

| Patella caerulea   | Croatia        | ISTPC1   | 0.9              | 10.1038/s41598-019-39959-9                                     |

@@

@@

>>>>

```{r Library and datasets}
#| echo: FALSE
#| message: FALSE
#| warning: FALSE

@@

# Scripts
# source("DTW_Script.R")
source("DTW_Script-update.R")

@@

Ferguson_data <- read_csv(here("data","Ferguson-data.csv")) %>% 
  na.omit()

@@

This paper is a short reassessment of Atlantic limpet shells and their correlations to Mg/Ca ratios.

In previous studies, various limpet specimens (Patella vulgata, Nacella magellanica and Nacella odifosdio) have been studied using Mg/Ca with mixed results. While Patella vulgata along the Spanish shoreline has since then repeatedly produced reliable correlations between sst and Mg/Ca ratios, this is not the case for other species.

In this study, we present elemental maps of various such species together with stable oxygen isotope values for some of the specimens. Some of these...

@@

# sample <- Samples[12]

@@

  temp_MgCa <- MgCa %>% 
    filter(sampleID == sample) %>% 
    filter(!(sampleID=="ND-1016-4"&dist<0.3)) %>%

    select(mg_ca) %>% 
    mutate(sample_seq = row_number())

@@

P_all <- Corr_data %>% 
  pivot_wider(values_from = "value",names_from = "proxy") %>% 
  mutate(region = substr(sampleID, 1, 1)) %>% 
  ggplot()+
  aes(d18o,mg_ca,col=sampleID)+
  geom_point(show.legend = FALSE)+
  labs(x=expression("‰ δ" ^ "18" * "O [VPDB]"),y="Mg/Ca intensity ratio")+
  stat_smooth(aes(fill = sampleID,col=sampleID), method = "lm", formula = y~x,show.legend = FALSE) +
  # stat_regline_equation(label.x = 2, #label.y = 32
  #   aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
  #   formula = y~x
  # ) +
  scale_colour_manual(values = c( "#33a02c","#2ecc71", "#ffc100", "#f39c12", "#007f7f", "#1f78b4", "#99d8c9", "#542788","black", "#d73027", "#EC7063","pink1")) +
scale_fill_manual(values = c( "#33a02c","#2ecc71", "#ffc100", "#f39c12", "#007f7f", "#1f78b4", "#99d8c9", "#542788","black", "#d73027", "#EC7063","pink1"))#+
  # facet_wrap(~region,ncol=2,scales = "free")

@@

# Discussion

```{r Ferguson Data}
#| fig-cap: Correlation graphs for Ferguson et al. specimens
#| fig-height: 20
#| fig-width: 10

Ferg_all <-
  Ferguson_data %>% 
  ggplot()+
  aes(d18o,mg_ca,col=shell)+
  geom_point(show.legend = FALSE)+
  labs(x=expression("‰ δ" ^ "18" * "O [VPDB]"),y="Mg/Ca intensity ratio")+
  stat_smooth(aes(fill = shell,col=shell), method = "lm", formula = y~x,show.legend = FALSE) +
  # stat_regline_equation(label.x = 2, #label.y = 32
  #   aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
  #   formula = y~x
  # ) +
  scale_colour_manual(values = c(  "#EC7063", "#007f7f", "#1f78b4", "#542788","black", "#d73027", "#EC7063")) +
scale_fill_manual(values = c(  "#EC7063", "#007f7f", "#1f78b4", "#542788","black", "#d73027", "#EC7063"))

Ferg_individual <- Ferg_all+
  facet_wrap(~shell,ncol=2)+#ylim(0.15,NA)+
  ggpubr::stat_regline_equation(label.x = 1, label.y = 26,
    aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),show.legend = FALSE,
    formula = y~x
  )+
  theme(
    strip.background = element_rect(fill = "white", colour = NA),
    strip.text = element_text(colour = "black"),
    panel.border = element_rect(colour="black")
  )

Ferg_all/Ferg_individual+plot_layout(heights = c(0.5,0.5))

@@

|-------------|-------------|-------------|-------------|----------------------|

@@

| *Patella caerulea* | Croatia        | ISTPC1   | 0.9              | 10.1038/s41598-019-39959-9                                                                                                         |

@@

| Patella rustica    | Gibraltar      | JL1      | 0.02             | [doi.org/10.1016/j.epsl.2011.05.054](https://doi.org/10.1016/j.epsl.2011.05.054 "Persistent link using digital object identifier") |
|                    |                | JL2      | 0.8 (0.79)       |                                                                                                                                    |
| Patella caerulea   | Gibraltar      | JM00     | 0.69 (0.79)      |                                                                                                                                    |
|                    |                | JM30     | 0.83 (0.79)      |                                                                                                                                    |

@@

## Reanalysis of ORK-LT5

```{r}

P1 <-
  MgCa %>% 
  filter(sampleID=="ORK_LT5") %>% 
  # filter(std<0.1) %>% 
  # filter(mg_ca<0.7) %>% 
  # filter(dist<20) %>% 
ggplot()+
  aes(dist,mg_ca)+
  geom_path(col="firebrick")+
  # xlim(0,10)+
  facet_wrap(~sampleID,scales = "free",nrow=1)

P2 <-
  D18O %>% 
  filter(sampleID=="ORK_LT5") %>% 
  ggplot()+
  aes(dist,d18O)+
  geom_path(col="cornflowerblue")+
  scale_y_reverse()+
  facet_wrap(~sampleID,scales = "free",nrow=1)

plot_compare <- P1/P2

plot_compare

```{r subsample}

# Function to create the desired subsampled tibble
subsample_data <- function(data, window_size = 0.05, step_size = 0.15) {
  max_dist <- max(data$dist)
  subsample_intervals <- seq(0, max_dist, by = step_size)

  subsampled_data <- subsample_intervals %>%
    purrr::map_df(~ {
      window_start <- .x
      window_end <- .x + window_size

      data_in_window <- data %>%
        filter(dist >= window_start & dist < window_end)

      if (nrow(data_in_window) > 0) {
        summarised_data <- data_in_window %>%
          summarise(across(c(x, y, z, mg_ca, std, relstd, dist), mean, .names = "mean_{.col}"))
        summarised_data <- summarised_data %>%
          mutate(window_start = window_start, window_end = window_end)
        return(summarised_data)
      } else {
        return(tibble())
      }
    })

  return(subsampled_data)
}

# Apply the function to your data
MgCa_filtered <- MgCa %>%
  filter(sampleID == "ORK_LT5")

MgCa_subsampled_150 <- subsample_data(MgCa_filtered,step_size = 0.15)

MgCa_subsampled_300 <-  subsample_data(MgCa_filtered,step_size = 0.30)

P3.1 <-
  MgCa_subsampled_150 %>% 
ggplot()+
  aes(mean_dist,mean_mg_ca)+
  geom_path(col="firebrick2")

P3.2 <-
  MgCa_subsampled_300 %>% 
ggplot()+
  aes(mean_dist,mean_mg_ca)+
  geom_path(col="firebrick4")

P1/P3.1/P3.2
